Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-04-26 products/manuscript/Manuscript.qmd

We will be looking for patterns in contraceptive methods and higher unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,short acting reversible, and traditional. Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies [@jonathanmarcbearakCountryspecificEstimatesUnintended2022]. According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group . Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study [@rivetContraceptionUseAttitudes2021]. African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies. This could be due to the current legal status of contraception within Latin American countries. Free emergency contraception access exists, but regulations differ. Chile, Colombia, and Ecuador legally recognize access, while Nicaragua and Bolivia rely on Ministerial protocols. Argentina and Brazil lack legal recognition, but offer protocols and guides. Mexico requires provision to victims of sexual and domestic violence [@heviaLegalStatusEmergency2012]. Stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging [@kavanaughContraceptiveMethodUse2018].

## Description of data and data source
The data source for this project is the Guttmacher Institute, which is a well-known research organization focusing on improving sexual and reproductive health around the world. The data used here is from their Adding It Up project. The goals of this project include estimating the need for, impact of, and costs associated with providing sexual and reproductive health services. Low and middle income countries are the target audience for this research, specifically women of reproductive age (15-49). Some of the variables included in this dataset are modern contraceptive use, unintended pregnancies, unplanned births, and abortions averted. Observations are drawn from nationally representative surveys, including Demographic and Health Surveys, UNICCEF Multiple Indicator Cluster Surveys, US Centers for Disease Control Reproductive Health Surveys, Performance Monitoring for Action Surveys, and others. The UN Population Division is the source of population projections for women in this age group, and estimates of unintended pregnancies are from the Guttmacher Institute, WHO, and other authors, which have been adjusted to 2019.

@@

Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among region?

>>>>

We will be looking for patterns in contraceptive methods and unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,short acting reversible, and traditional. Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies [@jonathanmarcbearakCountryspecificEstimatesUnintended2022]. According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group . Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study [@rivetContraceptionUseAttitudes2021]. African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies. This could be due to the current legal status of contraception within Latin American countries. Free emergency contraception access exists, but regulations differ. Chile, Colombia, and Ecuador legally recognize access, while Nicaragua and Bolivia rely on Ministerial protocols. Argentina and Brazil lack legal recognition, but offer protocols and guides. Mexico requires provision to victims of sexual and domestic violence [@heviaLegalStatusEmergency2012]. Stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging [@kavanaughContraceptiveMethodUse2018].

## Description of data and data source
The data source for this project is the Guttmacher Institute, which is a well-known research organization focusing on improving sexual and reproductive health around the world. The data used here is from their Adding It Up project. The goals of this project include estimating the need for, impact of, and costs associated with providing sexual and reproductive health services. Low and middle income countries are the target audience for this research, specifically women of reproductive age (15-49). Some of the variables included in this dataset are modern contraceptive use, unintended pregnancies, unplanned births, and abortions averted. Observations are drawn from nationally representative surveys, including Demographic and Health Surveys, UNICCEF Multiple Indicator Cluster Surveys, US Centers for Disease Control Reproductive Health Surveys, Performance Monitoring for Action Surveys, and others. The UN Population Division is the source of population projections for women in this age group, and estimates of unintended pregnancies are from the Guttmacher Institute, WHO, and other authors, which have been adjusted to 2019. It is important to note the types of contraception included in each category defined by the Adding It Up project. Permanent methods include female and male sterilization, and long acting methods include IUDs and implants. These methods are combined for several variables and include a 'larcster' prefix or suffix to identify them throughout the dataset. Short-acting methods include hormonal pills, injectables, male and female condoms, emergency contraceptive pills, patches, rings, diaphragms, vaginal spermicide, and other supply methods, which is denoted with a 'sarc' prefix or suffix in the variable name. Traditional methods include periodic abstinence, withdrawal, abstinence, and breast-feeding, which is denoted with a 'trad' prefix or suffix in the variable name.

@@

Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users between the five regions included in the Adding It Up project?
